The Tracing package\\ \large a more in depth explanation

### 2015-05-24 manual/tex/controller.tex
Update manual (added serialization + corrections)

The first part to take into consideration is the \textbf{allocator}, which decides on which simulation core to place which model. We have already included a simple allocator (called \textsl{SimpleAllocator}) that distributes models evenly, but for more complex behavior it is possible to create one tailored specifically for your simulation purposes by inheriting from the \textsl{Allocator} base class.\\

>>>>

The first part to take into consideration is the \textbf{allocator}, which decides on which simulation core to place which model. We have already included a simple allocator (called \textsl{SimpleAllocator}) that distributes models evenly, but for more complex behavior it is possible to create one tailored specifically to your simulation purposes by inheriting from the \textsl{Allocator} base class.\\

### 2015-05-24 manual/tex/faq.tex
Update manual (added serialization + corrections)

\section{Platform specific}
\subsubsection{Virtualization}
\begin{enumerate}
\item On Windows host, with a *nix image, it's possible that standard compliant code leaks memory in std::string. We're aware of this issue and resolve it where we detect it with Valgrind.
\item Threading is significantly slower depending on your host, you should at least allocate 2 virtual CPU's to your virtual instance to allow proper functioning of the simulator. Several threaded tests can fail , since the runtime can only provide a hint (std::thread::hardware\textunderscore concurrency()) to the program to detect if we can use threads.
\end{enumerate}
\subsubsection{Cygwin}
\begin{enumerate}
\item std::to\textunderscore string is not supported, do not use it in your own models.
\item GTest version 1.7 corrupts the stack if a test fails, use version 1.6 until GTest resolves this.
\end{enumerate}
\section{General considerations}
\begin{enumerate}
\item If you violate certain clearly documented preconditions (such as having several models with the exact same name), the simulator is allowed to crash hard.
\item Do not use any non-serializable features of C++, examples are std::function (function / raw pointers in general).
\item Time is represented using std::size\textunderscore t internally, which is 64 bits wide on most 64-bit systems, 32 on older 32-bit system. As a result, should you ever use an explicit value larger than std::numeric\textunderscore limits$<$std::size\textunderscore t$>$::max(), the time representation will fail in unexpected ways. The same issue arises if the simulator advances beyond that time point. It's possible to change the typedef t\textunderscore timestamp to use a larger type should this be necessary for you. For example :\\
\lstinline!typedef Time<unsigned long long, unsigned long long> t_timestamp;!
As long as the supplied type to Time has well-defined operators for the aritmetic and comparison operators (at least partial ordered), and the type has a specialization for numeric limits, it can be used.
\end{enumerate}
\section{Threading}
\begin{enumerate}
\item Never use more simulator cores than hardware threads you have available. For example on a dual core cpu with hyperthreading, 4 simualtor-cores is the limit you should adhere to. The simulator may work with more simulator-cores, but performance will be crippled and starvation is likely.
\item libstdc++'s implementation of std::string is Copy on Write, this introduces an inherent danger for race conditions. Should you ever be confronted with this issue (in your own models), use the provided copyString function in stringtools.h to force g++ to generate an explicit (safe) copy.
\end{enumerate}

>>>>

\section{General considerations}
\begin{enumerate}
	\item If you violate certain clearly documented preconditions (such as having several models with the exact same name), the simulator is allowed to crash hard.
	\item Do not use any non-serializable features of C++, examples are std::function (function / raw pointers in general).
	\item Time is represented using std::size\textunderscore t internally, which is 64 bits wide on most 64-bit systems, 32 on older 32-bit system. As a result, should you ever use an explicit value larger than std::numeric\textunderscore limits$<$std::size\textunderscore t$>$::max(), the time representation will fail in unexpected ways. The same issue arises if the simulator advances beyond that time point. It's possible to change the typedef t\textunderscore timestamp to use a larger type should this be necessary for you. For example :\\
	\lstinline!typedef Time<unsigned long long, unsigned long long> t_timestamp;!
	As long as the supplied type to Time has well-defined operators for the aritmetic and comparison operators (at least partial ordered), and the type has a specialization for numeric limits, it can be used.
\end{enumerate}

\section{Platform specific}
\subsubsection{Virtualization}
\begin{enumerate}
\item On Windows host, with a *nix image, it's possible that standard compliant code leaks memory in std::string. We're aware of this issue and resolve it where we detect it with Valgrind.
\item Threading is significantly slower depending on your host, you should at least allocate 2 virtual CPU's to your virtual instance to allow proper functioning of the simulator. Several threaded tests can fail , since the runtime can only provide a hint (std::thread::hardware\textunderscore concurrency()) to the program to detect if we can use threads.
\end{enumerate}
\subsubsection{Cygwin}
\begin{enumerate}
\item std::to\textunderscore string is not supported, do not use it in your own models. Some old C functions like atoi and stoi are not supported either.
\item GTest version 1.7 corrupts the stack if a test fails, use version 1.6 until GTest resolves this.
\end{enumerate}

\section{Threading}
\begin{enumerate}
\item Never use more simulator cores than hardware threads you have available. For example on a dual core cpu with hyperthreading, 4 simualtor cores is the limit you should adhere to. The simulator may work with more simulator cores, but performance will be crippled and starvation is likely.
\item libstdc++'s implementation of std::string is Copy on Write, this introduces an inherent danger for race conditions. Should you ever be confronted with this issue (in your own models), use the provided copyString function in stringtools.h to force g++ to generate an explicit (safe) copy.
\end{enumerate}

### 2015-05-24 manual/tex/models.tex
Update manual (added serialization + corrections)

\section{States}
You can can make your own states with almost no limitations. The only requirements for a user's own states are that they should be objects from a type that implements \texttt{n\_model::State}. This class has a few functions that have to be implemented:
\begin{itemize}
	\item \texttt{std::string toString()}
	\item \texttt{std::string toXML()}
	\item \texttt{std::string toCell()}
\end{itemize}
If the user does not override the above functions, this will result in a logged error. Serialization prohibits abstract functions, which enforces this workaround.\\
While members in sub-classes can be added without any problems, by default there is a public member \texttt{std::string m\_state} implemented in \texttt{State} that can be used to represent the user's state.
\\ For serialization you have to declare (after class definition) 
\begin{verbatim}
CEREAL_REGISTER_TYPE(MyState);
\end{verbatim}

\section{Atomic models}
Atomic models can be made by implementing a sub-class of \texttt{n\_model::AtomicModel}. The most basic functions to be implemented are:
\begin{enumerate}
	\item \texttt{void AtomicModel::extTransition(const std::vector{\textless}n\_network::t\_msgptr{\textgreater} \& inputs)}\\
		Performs the model's external transition. The input argument is a vector with all messages at a certain point in time.
	\item \texttt{void AtomicModel::intTransition()}\\
		Performs the model's internal transition. 
	\item \texttt{t\_timestamp AtomicModel::timeAdvance() const}\\
		Should return the model's current time advance value; the amount of time after which the following internal transition occurs. This function should not alter the model in any way.
	\item \texttt{std::vector{\textless}n\_network::t\_msgptr{\textgreater} AtomicModel::output()  const}\\
		Returns the model's output; all messages that are to be send should be returned via this function. This function should not alter the model in any way.
\end{enumerate}
If these are not overridden, this will result in a logged assertion error.\\
A user can optionally implement the lookahead function, when a parallel conservative simulation is required:
\begin{itemize}
	\item \texttt{t\_timestamp AtomicModel::lookAhead() const}
\end{itemize}

@@

When \textbf{sending messages} (\textbf{output function}), there are a few rules you should take into account. The message you want to send can only be a string. This string should be given to the port you want to send it through. The port will then create the actual messages (that are send over the models) and return them to you in a vector. All these messages (from possibly different ports) should be joined in 1 vector and be returned in the output function. The following functions will allow you to do so:
\begin{itemize}
	\item \texttt{t\_portptr Model::getPort(std::string name) const}
	\item \texttt{std::vector{\textless}n\_network::t\_msgptr{\textgreater} Port:: \\createMessages(std::string message)}
\end{itemize} 
Something like this would work: \texttt{this-{\textgreater}getPort("OUT")-{\textgreater} \\createMessages("my\_message");} \\

It's important that a model name is unique over the whole simulation, and portname should be unique per model.\\

For serialization you have to declare (after class definition)
\begin{verbatim}
CEREAL_REGISTER_TYPE(MyModel);
\end{verbatim}

@@

@@

>>>>

\section{States}
You can make your own states with almost no limitations. The only requirements for a user's own states are that they should be objects from a type that implements \texttt{n\_model::State}. This class has a few functions that have to be implemented:
\begin{itemize}
	\item \texttt{std::string toString()}
	\item \texttt{std::string toXML()}
	\item \texttt{std::string toCell()}
\end{itemize}
If the user does not override the above functions, this will result in a logged error. Serialization prohibits abstract functions, which enforces this workaround.\\
While members in sub-classes can be added without any problems, by default there is a public member \texttt{std::string m\_state} implemented in \texttt{State} that can be used to represent the user's state.

\subsubsection{Serialization}
For serialization you have to define three serialization functions with following signature: 
\begin{lstlisting}
void serialize(n_serialization::t_oarchive& archive) {
	archive(cereal::virtual_base_class<State>( this ));
}
\end{lstlisting}
\begin{lstlisting}
void serialize(n_serialization::t_iarchive& archive) {
	archive(cereal::virtual_base_class<State>( this ));
}
\end{lstlisting}
\begin{lstlisting}
static void load_and_construct(n_serialization::t_iarchive& archive,
			cereal::construct<MyState>& construct) {
	construct("");
	construct->serialize(archive);
}
\end{lstlisting}
In case your own states uses more datamembers, you can include them to the serialization by adding them at the back of the parameter list of archive() in the first two functions. For more information consult the Cereal manual.\\
\\ 
For example:
\begin{lstlisting}
int a = 5;
archive(cereal::virtual_base_class<State>( this ), a);
\end{lstlisting}
After the class definition you have to register your new state throug the following macro: 
\begin{lstlisting}
CEREAL_REGISTER_TYPE(MyState);
\end{lstlisting}
\textbf{NOTE} the registration has to be done outside of any namespaces to work properly.

\section{Atomic models}
Atomic models can be made by implementing a sub-class of \texttt{n\_model::AtomicModel}. The most basic functions to be implemented are:
\begin{enumerate}
	\item \texttt{void AtomicModel::extTransition(const std::vector{\textless}n\_network::t\_msgptr{\textgreater} \& inputs)}\\
		Performs the model's external transition. The input argument is a vector with all messages at a certain point in time.
	\item \texttt{void AtomicModel::intTransition()}\\
		Performs the model's internal transition. 
	\item \texttt{t\_timestamp AtomicModel::timeAdvance() const}\\
		Should return the model's current time advance value, the amount of time after which the following internal transition occurs. This function should not alter the model in any way.
	\item \texttt{std::vector{\textless}n\_network::t\_msgptr{\textgreater} AtomicModel::output()  const}\\
		Returns the model's output; all messages that are to be sent should be returned via this function. This function should not alter the model in any way.
\end{enumerate}
If these are not overridden, this will result in a logged assertion error.\\
A user can optionally implement the lookahead function, when a parallel conservative simulation is required:
\begin{itemize}
	\item \texttt{t\_timestamp AtomicModel::lookAhead() const}
\end{itemize}

@@

When \textbf{sending messages} (\textbf{output function}), there are a few rules you should take into account. The message you want to send can only be a string. This string should be given to the port you want to send it through. The port will then create the actual messages (which are sent over the models) and return them to you in a vector. All these messages (from possibly different ports) should be joined in 1 vector and be returned in the output function. The following functions will allow you to do so:
\begin{itemize}
	\item \texttt{t\_portptr Model::getPort(std::string name) const}
	\item \texttt{std::vector{\textless}n\_network::t\_msgptr{\textgreater} Port:: \\createMessages(std::string message)}
\end{itemize} 
Something like this would work: \texttt{this-{\textgreater}getPort("OUT")-{\textgreater} \\createMessages("my\_message");} \\

It's important that a model name is unique over the whole simulation, and port name should be unique per model. \\

@@

\subsubsection{Serialization}
For serialization you have to define three serialization methods with following signature: 
\begin{lstlisting}
void serialize(n_serialization::t_oarchive& archive) {
	archive(cereal::virtual_base_class<AtomicModel>( this ));
}
\end{lstlisting}
\begin{lstlisting}
void serialize(n_serialization::t_iarchive& archive) {
	archive(cereal::virtual_base_class<AtomicModel>( this ));
}
\end{lstlisting}
\begin{lstlisting}
static void load_and_construct(n_serialization::t_iarchive& archive,
			cereal::construct<MyAtomicModel>& construct) {
	construct("");
	construct->serialize(archive);
}
\end{lstlisting}
In case your own atomic models uses more datamembers, you can include them to the serialization by adding them at the back of the parameter list of archive() in the first two functions. For more information consult the Cereal manual.\\
\\ 
For example:
\begin{lstlisting}
int a = 5;
archive(cereal::virtual_base_class<AtomicModel>( this ), a);
\end{lstlisting}
After the class definition you have to register your new atomic model throug the following macro: 
\begin{lstlisting}
CEREAL_REGISTER_TYPE(MyAtomicModel);
\end{lstlisting}

\textbf{NOTE} the registration has to be done outside of any namespaces to work properly.

@@

\subsubsection{Serialization}
For serialization you have to define three serialization methods with following signature: 
\begin{lstlisting}
void serialize(n_serialization::t_oarchive& archive) {
	archive(cereal::virtual_base_class<CoupledModel>( this ));
}
\end{lstlisting}
\begin{lstlisting}
void serialize(n_serialization::t_iarchive& archive) {
	archive(cereal::virtual_base_class<CoupledModel>( this ));
}
\end{lstlisting}
\begin{lstlisting}
static void load_and_construct(n_serialization::t_iarchive& archive,
			cereal::construct<MyCoupledModel>& construct) {
	construct("");
	construct->serialize(archive);
}
\end{lstlisting}
In case your own coupled models uses more datamembers, you can include them to the serialization by adding them at the back of the parameter list of archive() in the first two functions. For more information consult the Cereal manual.\\
\\ 
For example:
\begin{lstlisting}
int a = 5;
archive(cereal::virtual_base_class<CoupledModel>( this ), a);
\end{lstlisting}
After the class definition you have to register your new atomic model throug the following macro: 
\begin{lstlisting}
CEREAL_REGISTER_TYPE(MyCoupledModel);
\end{lstlisting}

\textbf{NOTE} the registration has to be done outside of any namespaces to work properly.
